---
title: "A Comprehensive Guide to Combining R and Python with `reticulate`"
author: "Alejandro Leonardo García Navarro"
date: "2024-06-20"
output: 
  pdf_document:
    toc: true
urlcolor: blue
---

```{r setup, include=FALSE}
library(reticulate)
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
## What is `reticulate`?
The `reticulate` package is a tool that allows to combine R and Python. It allows users to call Python from R and R from Python, combining the strengths of both programming languages in a single workflow.

With this library, you can import any Python module and access its functions, classes, and objects from R, enabling a more versatile and flexible approach to data analysis, machine learning, and statistical computing.

## Benefits of combining R and Python
Combining R and Python brings together the best of both worlds:

1. Choose the best tool for each task by leveraging R's statistical analysis and Python's programming and machine learning strengths.
2. Access more libraries and packages from both ecosystems.
3. Easy transfer of data between R and Python for flexible data handling in complex analysis pipelines.

## Prerequisites and installation
Before using the library, make sure you have the following prerequisites:

1. R Installation: Make sure you have R installed on your system. You can download it from  [CRAN](https://cran.r-project.org/).
2. Python Installation: [Install Python](https://www.python.org/downloads/) on your system.
3. RStudio (Optional but recommended): Using RStudio as your IDE can simplify the process of using `reticulate`. Download RStudio from [here](https://posit.co/download/rstudio-desktop/).

Once you have completed all the prerequisites, it is time to install the package. Use the following command in your R console:
```{r warning=FALSE}
install.packages("reticulate")
```

After installation, load the package using:
```{r}
library(reticulate)
```



# Basic Usage




